### 2020-05-27 appendix.tex

\clearpage
\makeatletter
\efloat@restorefloats
\makeatother

\begin{appendix}
\section{}
Given the inherent issues found in event data sets (Wang, Kennedy,
Lazer, \& Ramakrishnan, 2016), we conducted the same analyses on the
Cline Center Historical Phoenix Event Data (Althaus, Bajjalieh, Carter,
Peyton, \& Shalmon, 2020). After filtering the data and creating the
count variables for all, positive, and negative events each day, there
was not enough variability to transform the data into deciles for the
target only Phoenix data. As a results, sextiles were used for these
analyses. For comparison, the filtered ICEWS data was also transformed
into sextiles.

\hypertarget{crqa-for-phoenix-data-in-sextiles}{%
\subsection{CRQA for Phoenix data in
sextiles}\label{crqa-for-phoenix-data-in-sextiles}}

\begin{longtable}[]{@{}cccccccccc@{}}
\caption{CRQA results for target only Phoenix data.}\tabularnewline
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative
events & p & sig.\tabularnewline
\midrule
\endfirsthead
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative
events & p & sig.\tabularnewline
\midrule
\endhead
DET & 31.065 & 0.241 & & 38.955 & 0.000 & *** & 31.854 & 0.139
&\tabularnewline
NRLINE & 144.000 & 0.229 & & 174.000 & 0.000 & *** & 148.000 & 0.105
&\tabularnewline
maxL & 4.000 & 0.421 & & 5.000 & 0.087 & . & 4.000 & 0.452
&\tabularnewline
L & 2.188 & 0.544 & & 2.270 & 0.039 & * & 2.182 & 0.580 &\tabularnewline
ENTR & 0.513 & 0.526 & & 0.650 & 0.033 & * & 0.504 & 0.550
&\tabularnewline
rENTR & 0.467 & 0.290 & & 0.469 & 0.276 & & 0.459 & 0.310
&\tabularnewline
LAM & 32.051 & 0.284 & & 46.154 & 0.008 & ** & 30.769 & 0.314
&\tabularnewline
TT & 2.273 & 0.196 & & 2.400 & 0.072 & . & 2.400 & 0.063 &
.\tabularnewline
\bottomrule
\end{longtable}

\begin{longtable}[]{@{}cccccccccc@{}}
\caption{CRQA results for source and target Phoenix
data.}\tabularnewline
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative
events & p & sig.\tabularnewline
\midrule
\endfirsthead
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative
events & p & sig.\tabularnewline
\midrule
\endhead
DET & 31.558 & 0.178 & & 37.377 & 0.000 & *** & 25.740 & 0.988
&\tabularnewline
NRLINE & 145.000 & 0.189 & & 169.000 & 0.000 & *** & 119.000 & 0.989
&\tabularnewline
maxL & 4.000 & 0.467 & & 5.000 & 0.084 & . & 4.000 & 0.454
&\tabularnewline
L & 2.207 & 0.368 & & 2.243 & 0.123 & & 2.193 & 0.509 &\tabularnewline
ENTR & 0.547 & 0.343 & & 0.610 & 0.104 & & 0.521 & 0.492
&\tabularnewline
rENTR & 0.498 & 0.135 & & 0.440 & 0.432 & & 0.474 & 0.244
&\tabularnewline
LAM & 38.462 & 0.073 & . & 48.718 & 0.000 & *** & 21.795 & 0.790
&\tabularnewline
TT & 2.500 & 0.017 & * & 2.235 & 0.280 & & 2.429 & 0.048 &
*\tabularnewline
\bottomrule
\end{longtable}

CRQ analyses for sextiled Phoenix event data found increased levels of
coupling between positive events and social cohesion, irrespective of
target or source and target filtering. Negative events were not coupled
with social cohesion irrepective of filtering.

\hypertarget{crqa-for-icews-data-in-sextiles}{%
\subsection{CRQA for ICEWS data in
sextiles}\label{crqa-for-icews-data-in-sextiles}}

\begin{longtable}[]{@{}cccccccccc@{}}
\caption{CRQA results for target only ICEWS data.}\tabularnewline
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative
events & p & sig.\tabularnewline
\midrule
\endfirsthead
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative
events & p & sig.\tabularnewline
\midrule
\endhead
DET & 39.250 & 0.000 & *** & 37.870 & 0.000 & *** & 38.659 & 0.000 &
***\tabularnewline
NRLINE & 177.000 & 0.000 & *** & 169.000 & 0.000 & *** & 170.000 & 0.000
& ***\tabularnewline
maxL & 5.000 & 0.093 & . & 5.000 & 0.081 & . & 5.000 & 0.087 &
.\tabularnewline
L & 2.249 & 0.089 & . & 2.272 & 0.033 & * & 2.306 & 0.009 &
**\tabularnewline
ENTR & 0.619 & 0.074 & . & 0.653 & 0.030 & * & 0.705 & 0.008 &
**\tabularnewline
rENTR & 0.447 & 0.368 & & 0.471 & 0.261 & & 0.509 & 0.118
&\tabularnewline
LAM & 43.590 & 0.013 & * & 39.744 & 0.047 & * & 42.308 & 0.030 &
*\tabularnewline
TT & 2.267 & 0.208 & & 2.385 & 0.065 & . & 2.357 & 0.092 &
.\tabularnewline
\bottomrule
\end{longtable}

\begin{longtable}[]{@{}cccccccccc@{}}
\caption{CRQA results for source and target ICEWS data.}\tabularnewline
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative
events & p & sig.\tabularnewline
\midrule
\endfirsthead
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative
events & p & sig.\tabularnewline
\midrule
\endhead
DET & 34.122 & 0.014 & * & 37.771 & 0.000 & *** & 39.448 & 0.000 &
***\tabularnewline
NRLINE & 152.000 & 0.064 & . & 168.000 & 0.000 & *** & 173.000 & 0.000 &
***\tabularnewline
maxL & 5.000 & 0.092 & . & 5.000 & 0.093 & . & 7.000 & 0.002 &
**\tabularnewline
L & 2.276 & 0.030 & * & 2.280 & 0.024 & * & 2.312 & 0.003 &
**\tabularnewline
ENTR & 0.663 & 0.022 & * & 0.661 & 0.022 & * & 0.698 & 0.005 &
**\tabularnewline
rENTR & 0.478 & 0.211 & & 0.477 & 0.242 & & 0.434 & 0.475
&\tabularnewline
LAM & 34.615 & 0.178 & & 39.744 & 0.047 & * & 48.718 & 0.002 &
**\tabularnewline
TT & 2.250 & 0.233 & & 2.385 & 0.067 & . & 2.235 & 0.295
&\tabularnewline
\bottomrule
\end{longtable}

CRQ analyses for sextiled ICEWS event data found coupling between both
positive and negative events with social cohesion, irrespective of
target or source and target filtering. Negative events were more stongly
coupled with social cohesion than positive events for the target only
analyses. The abscence of coupling between negative events and social
cohesion in the Phoenix CRQ analyses speaks to the unreliability of
event data, and that nonlinear analyses in general are only as valuable
as the quality of the data being used.

\hypertarget{windowed-crqa-for-phoenix-data-in-sextiles}{%
\subsection{Windowed CRQA for Phoenix data in
sextiles}\label{windowed-crqa-for-phoenix-data-in-sextiles}}

\hypertarget{syria-as-target}{%
\subsubsection{Syria as Target}\label{syria-as-target}}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/Phoenix/windowed-crqa/target/windowed_all_RR} \caption{Windowed CRQA visualization of RR for social cohesion and count of all events from target filtered data.}(\#fig:plot-RR-targ-all-secondary-Phoenix)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/Phoenix/windowed-crqa/target/windowed_all_DET} \caption{Windowed CRQA visualization of DET for social cohesion and count of all events from target filtered data.}(\#fig:plot-DET-targ-all-secondary-Phoenix)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/Phoenix/windowed-crqa/target/windowed_pos_RR} \caption{Windowed CRQA visualization of RR for social cohesion and count of positive events from target filtered data.}(\#fig:plot-RR-targ-pos-secondary-Phoenix)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/Phoenix/windowed-crqa/target/windowed_pos_DET} \caption{Windowed CRQA visualization of DET for social cohesion and count of positive events from target filtered data.}(\#fig:plot-DET-targ-pos-secondary-Phoenix)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/Phoenix/windowed-crqa/target/windowed_neg_RR} \caption{Windowed CRQA visualization of RR for social cohesion and count of negative events from target filtered data.}(\#fig:plot-RR-targ-neg-secondary-Phoenix)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/Phoenix/windowed-crqa/target/windowed_neg_DET} \caption{Windowed CRQA visualization of DET for social cohesion and count of negative events from target filtered data.}(\#fig:plot-DET-targ-neg-secondary-Phoenix)
\end{figure}

\hypertarget{syria-as-source-and-target}{%
\subsubsection{Syria as Source and
Target}\label{syria-as-source-and-target}}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/Phoenix/windowed-crqa/source_target/windowed_all_RR} \caption{Windowed CRQA visualization of RR for social cohesion and count of all events from source and target filtered data.}(\#fig:plot-RR-source-targ-all-secondary-Phoenix)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/Phoenix/windowed-crqa/source_target/windowed_all_DET} \caption{Windowed CRQA visualization of DET for social cohesion and count of all events from source and target filtered data.}(\#fig:plot-DET-source-targ-all-secondary-Phoenix)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/Phoenix/windowed-crqa/source_target/windowed_pos_RR} \caption{Windowed CRQA visualization of RR for social cohesion and count of positive events from source and target filtered data.}(\#fig:plot-RR-source-targ-pos-secondary-Phoenix)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/Phoenix/windowed-crqa/source_target/windowed_pos_DET} \caption{Windowed CRQA visualization of DET for social cohesion and count of positive events from source and target filtered data.}(\#fig:plot-DET-source-targ-pos-secondary-Phoenix)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/Phoenix/windowed-crqa/source_target/windowed_neg_RR} \caption{Windowed CRQA visualization of RR for social cohesion and count of negative events from source and target filtered data.}(\#fig:plot-RR-source-targ-neg-secondary-Phoenix)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/Phoenix/windowed-crqa/source_target/windowed_neg_DET} \caption{Windowed CRQA visualization of DET for social cohesion and count of negative events from source and target filtered data.}(\#fig:plot-DET-source-targ-neg-secondary-Phoenix)
\end{figure}

\hypertarget{windowed-crqa-for-icews-data-in-sextiles}{%
\subsection{Windowed CRQA for ICEWS data in
sextiles}\label{windowed-crqa-for-icews-data-in-sextiles}}

\hypertarget{syria-as-target-1}{%
\subsubsection{Syria as Target}\label{syria-as-target-1}}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/icews/windowed-crqa/target/windowed_all_RR} \caption{Windowed CRQA visualization of RR for social cohesion and count of all events from target filtered data.}(\#fig:plot-RR-targ-all-secondary-icews)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/icews/windowed-crqa/target/windowed_all_DET} \caption{Windowed CRQA visualization of DET for social cohesion and count of all events from target filtered data.}(\#fig:plot-DET-targ-all-secondary-icews)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/icews/windowed-crqa/target/windowed_pos_RR} \caption{Windowed CRQA visualization of RR for social cohesion and count of positive events from target filtered data.}(\#fig:plot-RR-targ-pos-secondary-icews)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/icews/windowed-crqa/target/windowed_pos_DET} \caption{Windowed CRQA visualization of DET for social cohesion and count of positive events from target filtered data.}(\#fig:plot-DET-targ-pos-secondary-icews)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/icews/windowed-crqa/target/windowed_neg_RR} \caption{Windowed CRQA visualization of RR for social cohesion and count of negative events from target filtered data.}(\#fig:plot-RR-targ-neg-secondary-icews)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/icews/windowed-crqa/target/windowed_neg_DET} \caption{Windowed CRQA visualization of DET for social cohesion and count of negative events from target filtered data.}(\#fig:plot-DET-targ-neg-secondary-icews)
\end{figure}

\hypertarget{syria-as-source-and-target-1}{%
\subsubsection{Syria as Source and
Target}\label{syria-as-source-and-target-1}}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/icews/windowed-crqa/source_target/windowed_all_RR} \caption{Windowed CRQA visualization of RR for social cohesion and count of all events from source and target filtered data.}(\#fig:plot-RR-source-targ-all-secondary-icews)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/icews/windowed-crqa/source_target/windowed_all_DET} \caption{Windowed CRQA visualization of DET for social cohesion and count of all events from source and target filtered data.}(\#fig:plot-DET-source-targ-all-secondary-icews)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/icews/windowed-crqa/source_target/windowed_pos_RR} \caption{Windowed CRQA visualization of RR for social cohesion and count of positive events from source and target filtered data.}(\#fig:plot-RR-source-targ-pos-secondary-icews)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/icews/windowed-crqa/source_target/windowed_pos_DET} \caption{Windowed CRQA visualization of DET for social cohesion and count of positive events from source and target filtered data.}(\#fig:plot-DET-source-targ-pos-secondary-icews)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/icews/windowed-crqa/source_target/windowed_neg_RR} \caption{Windowed CRQA visualization of RR for social cohesion and count of negative events from source and target filtered data.}(\#fig:plot-RR-source-targ-neg-secondary-icews)
\end{figure}

\begin{figure}[!h]
\includegraphics[width=4in,]{./results/secondary/icews/windowed-crqa/source_target/windowed_neg_DET} \caption{Windowed CRQA visualization of DET for social cohesion and count of negative events from source and target filtered data.}(\#fig:plot-DET-source-targ-neg-secondary-icews)
\end{figure}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-illinoisdatabankIDB-2796521}{}%
Althaus, S., Bajjalieh, J., Carter, J., Peyton, B., \& Shalmon, D.
(2020). \emph{Cline center historical phoenix event data. Cline center
for advanced social research. V1.3.0. May 4} {[}Data set{]}. University
of Illinois at Urbana-Champaign.
\url{https://doi.org/10.13012/B2IDB-0647142_V3}

\leavevmode\hypertarget{ref-wang2016growing}{}%
Wang, W., Kennedy, R., Lazer, D., \& Ramakrishnan, N. (2016). Growing
pains for global monitoring of societal events. \emph{Science},
\emph{353}(6307), 1502--1503.
\end{appendix}