### 2020-06-01 manuscript.Rmd
Updates to Lexical Cohesion section and Cohesion structure

---
title             : "Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring"
shorttitle        : "Arab Spring"

author:
  - name          : "Megan Chiovaro"
    affiliation   : "1, 2"
    corresponding : yes    # Define only one corresponding author
    address       : "406 Babbidge Road, Unit 1020, Storrs, CT 06269"
    email         : "megan.chiovaro@uconn.edu"
  - name          : "Alexandra Paxton"
    affiliation   : "1,2,*"
  - name          : "Leah C. Windsor"
    affiliation   : "3,*"

affiliation:
  - id            : "1"
    institution   : "Center for the Ecological Study of Perception and Action, University of Connecticut"
  - id            : "2"
    institution   : "Department of Psychological Sciences, University of Connecticut"
  - id            : "3"
    institution   : "University of Memphis"
  - id            : "*"
    institution   : "Equal author contributions"

abstract: |
  Social mobilization requires effective collective coordination. In recent 
  years, individuals involved in social movements have taken to social media
  platforms to reach broader audiences at rapid speeds. Effective virtual 
  mobilization similarly requires social cohesion and coordination, yet the 
  degree to which virtual mobilization reflects or drives real-world movements
  is unclear. Here, we explore the dynamics of real-world events and social 
  cohesion on Twitter in Syria during the Arab Spring. Using the nonlinear 
  methods cross-recurrence quantification analysis and windowed cross-recurrence
  quantification analysis, we investigate if counts of all daily events, daily 
  positive events, and daily negative events are coupled with social cohesion 
  found in Syrian tweets. Results indicate that virtual social cohesion is 
  coupled with the counts of all and negative events each day but shows a more
  nuanced connection to positive events, only visible when examining finer-grained
  data. This suggests that virtual mobilization does reflect 
  real-world movements, with the strongest connections to the frequency of all
  events and the frequency of negative events. We 
  conclude with a discussion of implications and applications of nonlinear 
  methods in political science research.
  <!-- https://tinyurl.com/ybremelq -->

keywords          : "social cohesion, recurrence quantification analysis, complex systems, Twitter, Arab Spring"
wordcount         : "X"

bibliography      : ["references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
appendix          :

documentclass     : "apa6"
classoption       : "man"
header-includes   :
  - "\\usepackage{float}"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}

# clear the workspace
rm(list=ls())

# install and load required packages
source('./scripts/required_packages.r')
source('./scripts/libraries_and_functions.r')

# turn on caching
knitr::opts_chunk$set(cache=TRUE, autodep=TRUE, cache.lazy=FALSE)

# create subdirectories
dir.create("./data/raw") # load three data files here
dir.create("./data/formatted/primary", recursive = TRUE)
dir.create("./data/formatted/secondary/icews", recursive = TRUE)
dir.create("./data/formatted/secondary/phoenix", recursive = TRUE)
dir.create("./results/primary/crqa", recursive = TRUE)
dir.create("./results/primary/windowed-crqa", recursive = TRUE)
dir.create("./results/secondary/icews/crqa", recursive = TRUE)
dir.create("./results/secondary/icews/windowed-crqa", recursive = TRUE)
dir.create("./results/secondary/phoenix/crqa", recursive = TRUE)
dir.create("./results/secondary/phoenix/windowed-crqa", recursive = TRUE)

```

```{r data-prep, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, results='hide'}

# format the data
source('./scripts/primary-analyses/01_formatting_data.R')
source('./scripts/secondary-analyses/icews/01_formatting_data.R')
source('./scripts/secondary-analyses/phoenix/01_formatting_data.R')

```

Prior to the advent of social media, people coordinated their protest behavior
using phone tree calling sheets and mimeographed flyers,
through in-person meetings of social organizations and churches, and through
word of mouth. They gathered together and hand-made protest signs that reflected
their agendas. As McAdam found, people were more likely to join movements when
they had a personal connection [@mcadam1986recruitment; @mcadam1993specifying]: 
Historically, strong ties have been
strong predictors of involvement in high-risk protest activity.

The cohesiveness and strength of social movements in the pre-Internet era could be
operationalized by the number of people turning out to protest and through the
political events that precipitated and resulted from social gatherings *en masse*.
The content of placards and protest
signs provided the protest narrative in their own words. Cohesive
social movements were reflected in their cohesive placards and posters. For
example, the ``I am a man'' posters held by the men involved in the 1968
sanitation workers' strike reflected their internal cohesion and unified,
goal-oriented movement [@Windsor2020].

However, the rise of the Internet and its decentralized online networks
has upended theories about recruitment to---and participation in---social
movements. Strong interpersonal ties have yielded to weak impersonal ties that
strongly predict involvement in social movements [@steinert2017spontaneous]. Research
methodologies for measuring social movements have similarly evolved from
ethnographic interviews to ``big data'' analyses of language on social media.

Today, social media replicates and amplifies
this phenomenon during moments of social unrest. Amid the cacophony of messages
during political upheaval---such as the Arab Spring---consistent themes emerge
through hashtags, retweets, and the repeated mentions of concepts such as
regime, security, demonstration, and human rights. The cohesiveness of these
themes (i.e., whether movement participants are metaphorically singing in one voice) 
reflects the real-world events happening contemporaneously with collective action in
social movements or civil conflicts.

In the present research, we investigate the relationship between linguistic
cohesion and social cohesion in times of social conflict and unrest. We tackle a
question related to the one posed by Zeitzoff [-@zeitzoff2017social]: How does 
social media reflect on-the-ground realities during times of social conflict?
To do so, we examine patterns of linguistic cohesion on Twitter and their
connection to real-world events during
the Arab Spring uprising in Syria between March and June 2012.

Disentangling the complicated relationship between mass online mobilization---and
social mobilization in person---continues to be relevant, even as our modes
of communication change. The
political processes that transpired during and after the Arab Spring are still
in the process of being understood. As Zeitzoff [-@zeitzoff2017social] recognizes, 
social media is changing conflict, but the mechanisms and relationships by
which it does so are still being understood.

## Syria in Context
The decades preceding the Arab Spring in Syria were marked by
social repression and fear. In 1971, Hafez al-Assad overthrew the previous
leader and remained in power for thirty years, presiding during
regional contentiousness and cooperation until his death in 2000. His son,
Bashar al-Assad continued the regime of authoritarian leadership as he assumed
his office. The Assad lineage hails from a minority Islamic sect (the Alawites)
in a multi-religious and multi-ethnic country comprising Christians, Sunni and
Shia Muslims, Kurds, and Palestinian refugees.

In December 2010, a Tunisian man named Mohamed Bouazizi was harassed and
unjustifiably stripped of his vendor cart by municipal officials. His
humiliation led him to set himself on fire in protest to the authoritarian rule,
ultimately leading to his death several days later. His actions served as a
catalyst for the Tunisian Revolution, and one week after his death, Zine El
Abidine Ben Ali---the longstanding authoritarian president---fled the country.
The events of Tunisia rapidly spread to surrounding countries also subject to
unjust authoritarian rule, including Yemen, Egypt, Libya, and Syria.

Following a wave of democratization efforts across the Middle East and North
Africa, modest Syrian protests began in March 2011. The regime responded
harshly with imprisonments, detentions, censorship, and military operations as
the resistance to Assad’s regime grew and spread throughout the country. The
fighting escalated dramatically in March 2012, after increased unrest in the
western city of Homs. International actors---including the Red Cross/Red Crescent
Society, the United Nations General Assembly, the Security Council, and
individual world leaders---worked toward cease-fires, negotiations,
and concessions without much success. While many high-level military and
diplomatic officials defected from the regime, numerous loyal forces
continued the regime's repression, and the Syrian people turned to social
media to help coordinate their activism and communicate to the world. Between March and
June 2012, more than 8 million tweets were sent within Syria, revealing the
details of the conflict suppressed by the government.

Syria is a multi-ethnic, multi-lingual, and multi-religious state. Kurdish
people make up between 10-15% of the population, Alawites between 8-15%,
Christians about 10%. There are also much smaller sects (e.g., Ismailis,
Druze), but Sunni Muslims comprise the vast majority of society. Most people
speak Arabic as their primary language. Most policy reports addressing the current
conflict report on the cohesion of the ruling Alawite military elite, rather
than on the cohesiveness of the opposition movement. However, the Sunni
majority's disaffection with the Syrian government and their lack of political
representation provide sufficient grievance and collective group identification
to make them a cohesive force.

## Social Media and Social Movements
During the Arab Spring, social media provided real-time information about the
events transpiring in countries across the Middle East and North Africa. Our
understanding about the role of social media in social movements is still
evolving. Some argue that these mass movements would not have been possible
without efficient large-scale coordination, both physically on the ground and
virtually on social media platforms.

Along with Facebook, Twitter has been
credited as essential to the Arab Spring---both for protesters (as a means to
coordinate with one another and to communicate with the world) and for the
government response (to monitor and track protestors to crack down on dissent)
[@comninos2011twitter]. The events of Arab Spring have been thoroughly discussed
within the political science literature [e.g., @lutterbeck2013arab;
@koehler2017political], but to our knowledge, the direct *coupling* between
Syrian Twitter cohesion and international action coordination has not been
investigated. In fact, some have been skeptical about the effect of social media on
collective action at all [@seay2014slacktivism].

This skepticism can be summarized in a quote by Lisa Anderson:

> The important story about the 2011
Arab revolts in Tunisia, Egypt, and Libya is not how the globalization of the
norms of civic engagement shaped the protesters' aspirations. Nor is it about
how activists used technology to share ideas and tactics. Instead, the critical
issue is how and why these ambitions and techniques resonated in their various
local contexts. [anderson2011demystifying]

Others have similarly expressed hesitation about the role of online
organizing. Some have argued that real-life strong ties outweigh online weak ties and
that the power of the strong ties through in-person contact is critical
to recruiting people for higher-risk activities 
[@gladwell2010revolution; @mcadam1986recruitment].

Interestingly, these views may
actually help to make the case for bridging the gap between social media and
social mobilization. Such views suggest the strength of social ties in an online
environment is inextricably linked to the *context* and *ambitions* of the actors.
Moreover, research suggests that weak ties still retain considerable political
power: While strong ties tend to bind smaller groups, weak ties
tend to bind groups across distances [@granovetter1977strength].

Social media generates critical information about events, including whether and how
a demonstration will take place, how many
people participate, and how those in power (e.g., state, military, security apparatus)
respond. Similarly, these media platforms alert us to more global shifts within
a conflict, such as when fighting moves to a new city or region, the number of
casualties and wounded, and the types of weaponry used by parties to the
conflict [@zeitzoff2017social]. Social media is not entirely event-driven, however.
In monitoring social media channels, government forces may respond to the
perception that citizens are gaining strength and sympathy. Governments have
been shown to be quite tolerant of dissent so long as it does not metamorphose
into collective action [@king2013censorship].

Citizens communicate with each other with increased transparency due to the
ubiquitous availability of social media such as low-cost and egalitarian
platforms like Twitter. Citizens are able to access these tools via mobile
phones and computers with Internet connectivity, providing a stream of real-time
shared information. Citizens can share the location of events like protests and military
incursions, the degree of state repression, and the number of casualties. In
turn, the government---interested in remaining in power---monitors social media
platforms to ascertain emerging threats [@comninos2011twitter], 
such as where new groups are forming
and where existing groups are strengthening, where gatherings are held, and who
the idea entrepreneurs are. An example of a leader 
expressing his interest in
such public conversation is given by Bashar al-Assad in a speech on June 3, 2012:

>I know that I have been away from the media for a long time, and I have missed
having direct contact with the citizens, but I have always been following up
with the daily occurrences and gathering the information so that my speech can
be built on what is said by the street.

With few exceptions, Assad chose to keep social media platforms up and running.
Gohdes [-@gohdes2015pulling] notes that regimes can attempt to thwart the growth of social
protest online and in the streets by shutting down the Internet.
During these times, however, she finds that military repression increases.
However, government information about the citizens decreases; as such, many
regimes have elected to maintain Internet connectivity as a means of monitoring
and surveilling their citizens.

At the same time, the insulation
of online social platforms allows researchers to detail real-world social dynamics
at unprecedented resolution and scale [@paxton2017finding]. The digital traces left by
individuals and collectively as groups are regarded as a valuable source of
information that can offer powerful insights into states of social
disequilibrium, uprisings, and escalations into civil wars.
By examining these conversations, researchers can study the degree of cohesion between
speakers and the alignment between that cohesion and their collective political
behavior.

## Cohesion: A Multi-Leveled, Multi-Layered Concept

We contextualize linguistic cohesion with social mobilization and response to the relative
intensity of the conflict. Cohesion is a multi-layered concept referring to the
connectedness of words, thoughts, speakers, and actions. Cohesion reflects the
continuity of language and expression, and it shows the connectivity of
participants in a social movement. To address the relationship between language
and discourse and social movements, we adopt an interdisciplinary approach that
incorporates linguistics, psychology, and political science using nonlinear
dynamical systems methodology to model the relationship between language and
action.

### Lexical Cohesion: Connectedness of Words

The concept of lexical, or semantic, cohesion has already been applied to the
analysis of political speech (for example, the language of Margaret Thatcher) 
and non-political speech [@barzilay1999using; @morris1991lexical].

The words that comprise social media posts are the first level at which cohesion
occurs. In Figure \@ref(fig:cohesion-in-tweets), we see four sequential tweets, 
each containing four words.
Each sequential tweet contains some content included in the previous tweets. The
linguistic construct of content word overlap measures the relationship across
units of meaning---here, tweets---where identical and semantically similar words
are repeated [@graesser2004coh]. These repeated lexical
items constitute a theme; on Twitter, this might be a particular hashtag, a
name, a location, or other nominal category. This localized cohesion is
referential; words are repeated and co-referred across tweets, creating a sense
of continuity and cohesiveness of meaning and message.

Typically, these analyses focus on *content words* (e.g., nouns, adjectives,
verbs), which convey more meaningful semantic information than *function words*
(e.g., articles, conjunctions). While function words can provide important
information about complexity and language structure
[@tausczik2010psychological], content words have more specialized meaning and
are therefore more useful when researchers want to understand *what* people are
discussing, rather than *how* they are discussing it [@graesser2004coh;
@mcnamara2014automated].

In estimating cohesion in Twitter posts, tweets that share content words 
will yield higher cohesion values, even among speakers with opposing
views on the subject matter. Repetition of terms and concepts is captured by
reiteration, which is a type of lexical cohesion in which a subsequent item
repeats a previous item, or refers to it through synonymous language
[@halliday1976cohesion]. In social media, we can conceive of retweets (i.e., reposts
or forwards of previous messages) as a form of reiteration
that increase message cohesion. Since retweets replicate verbatim some part of
another tweet, these tweets would also have higher cohesion. Hashtags function
similarly and are included in our analysis.

```{r cohesion-in-tweets, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = "Add more informative caption."}

# call figure 1 here

```

## Interlocutor cohesion
Lexical cohesion contributes to the next level of relatedness among tweets:
interlocutor cohesion. This is the degree to which people in a conversation,
whether in person or online, have aligned their language [@Coupland_Giles_1988].
Known as communication accommodation theory, the degree of alignment between
speakers reveals their closeness, status, and deliberative strategy
[@Giles_2016]. Speakers can converge and diverge in their communication.
Convergence facilitates easier transmission of information, because the speakers
– or interlocutors – presume some existing familiarity or knowledge base between
them. When you know someone well enough to finish their sentence, this is an
example of convergence. Divergence, on the other hand, is more laborious because
the speakers must make their presumptions explicit and tacit. Interlocutor
cohesion, or convergence, happens when speakers are “singing in the same key” or
are “on the same page.” Windsor (2020) describes various types of cohesion,
using the example of the Occupy Wall Street movement as a symbol of divergence
given the myriad signs, slogans, and interests of the participants
[@Windsor_2020]. On the other hand, the “I am a man” protest captured by Ernest
Withers during the 1968 santiation workers’ strike in Memphis, TN, or the “Black
Lives Matter” placards following the uprising in Ferguson, MO, symbolize
convergence and strong interlocutor cohesion [@Edgar_Johnson_2018].

Cognitive and social psychology theories of alignment and coordination lend
themselves nicely to the study of social media communication during times of
conflict and contentious behavior [@Tollefsen_Dale_Paxton_2013;
@Garrod_Pickering_2004; @Pickering_Garrod_2004]. Over the past few decades, this
area has garnered significant attention and produced evidence showing that
people sometimes linguistically and behaviorally align as they interact more
[@Giles_Coupland_Coupland_1991; @Tollefsen_Dale_Paxton_2013]. Similarly,
accommodation theory refers to the ways in which individuals and groups
linguistically adapt to minimize the social space distance. Speakers in a
collective setting pursue strategies which increase their common ground, and
this is reflected in the language they use to communicate with one another.

A similar process happens among participants in social media discussions. The
process of speakers’ (or tweeters) language becoming more similar is called
linguistic style matching (LSM) [@Niederhoffer_Pennebaker_2002]. LSM is a form
of synchrony between speakers, akin to linguistic convergence. As Bernieri and
Rosenthal (1991) describe, people engaged in linguistic synchrony match each
others’ verbal behaviors, adopting not only similar terms, but similar patterns
of language [@Bernieri_Rosenthal_1991]. For example, LSM is particularly evident
when one speaker introduces swearing into a genteel conversation; the first
usage opens the vulgarity floodgates, and subsequent speakers may repeat the
initial curse, as well as innovate new curses into the conversation. Fed by
lexical cohesion, interlocutor cohesion ebbs and flows along a timeline
contextualized by variation in real world events. Interlocutor cohesion may
increase preceding an in-person social mobilization event, alerting a repressive
government that collective mobilization is happening. Interlocutor cohesion may
also respond to exogenous events such as spontaneous, unprovoked militarized
conflict. If government forces target a new location using military weapons,
this can precipitate online mobilization and discussion of the event.

## Social and Physical Cohesion
In turn, other studies show that linguistic alignment leads to increased social
connectedness and solidarity [@Miles_Nind_Macrae_2009], which is associated with
increased motivation for participation [@Walton_Cohen_Cwir_Spencer_2012],
facilitating successful coordination of group actions toward shared goals
[@Fusaroli_Bahrami_Olsen_Roepstorff_Rees_Frith_Tylén_2012;
@Richardson_Dale_2005]. This research emphasizes both the importance of and
connections between linguistic alignment and behavioral alignment. The more
people interact, the more similar their language and linguistic patterns become.
This holds true for face-to-face interactions as well as online communication
[@Riordan_Dale_Kreuz_Olney_2011}. Given this, we should expect to see increased
linguistic cohesion amongst people participating in online discussions about an
idea, event, or individual. Given that social media reports in real time about
actual events taking place, i.e., protests or demonstrations in which people are
assembled together in a location, or attacks that require coordination between
members of a unit, the cohesion in social media should be expected to correlate
with the social cohesion of the group [@Zeitzoff_2011}.

A socially cohesive group is inherently more threatening to the existing regime
because it reflects the professionalization of potential alternate leadership,
and collective opposition momentum. Groups with high social cohesion are more
difficult to divide and conquer, and more cohesive messages may also garner
increased international support. A paper published in Nature argues that global
insurgency movements all exhibit a similar pattern to their social structure
[@Bohorquez_Gourley_Dixon_Spagat_Johnson_2009}. The authors assert that each
violent group has a stable state, and that exogenous events can create
perturbations that push the group to become more cohesive, or alternatively to
become more fragmented. Rather than frenetic, patchwork messages via Twitter,
cohesive messages indicate increased organizational structure, which should
incentivize a response from the government. What process is the driving force,
though? Does an increase in linguistic cohesion precede or follow increased
repression by the government? Do movements get stronger first, and then the
government responds, or does the government increase its offensive measures, and
thus social groups have a clear opponent around whom they can coalesce both
socially and linguistically?

Thus, verbal synchrony leads to physical synchrony. Shared language in an online
environment can influence individuals’ propensity to engage in similar
coordinated behavior. In the case of the Arab Spring, this may mean social
mobilization in public spaces to protest against government brutality. Words in
a passage of text – or series of tweets – have some degree of semantic relation.
Individuals engaged in a conversation, whether in real life or online, have the
tendency to use similar language as they interact, increasing the cohesion of
their language. As an extension, multiple individuals collectively communicating
can also demonstrate linguistic alignment. In the context of social conflict,
the linguistic cohesion of participants in the online conversation should
correspond to the degree of social cohesion on the ground. We link these
concepts of social and linguistic cohesion together to demonstrate how language
serves as an indicator of the degree of social mobilization, using Twitter feeds
during the spring of 2012 in Syria.

Real-world collective action is known to bring about a sense of togetherness,
also known as “collective effervescence” [@durkheim1995elementary]. This feeling
of being a part of something larger than oneself leads individuals to act in
ways that would not have on their own. Being overcome with this feeling can lead
to violence [@collins2009violence; @spaaij2014sports]. Conversely, increased
exposure to violence can result in positive social cohesion. @gilligan2014civil
found that exposure to fatal wartime violence resulted in increased prosocial
motivations, including cooperative behaviors, feelings of trust, and overall
community-level social cohesion.

Physical and social cohesion in the world can be modeled using event data which
encodes who does what to whom on a given day. This could include events such as
the government (GOV) imposing a blockade or fighting with artillery or tanks on
a civilian group (CIV). CAMEO (Conflict and Mediation Event Observations) encode
a set of activities with numerical values, rescaled to reflect a
conflict-cooperation scale [@Schrodt_Hall_2006; @Goldstein_1992]. These event
codes include a range of behaviors, from making public statements and appeals,
to demanding, threatening, protesting, and using unconventional mass violence.
Steinert-Threlkeld (2017) modeled a similar relationship between social media
and social mobilization using a Gini coefficient measure, rather than a tf-idf
(term frequency-inverse document frequency) measure alongside event data
[@Ramos_2003; @Threlkeld_2017]. This pathbreaking work demonstrates that
participants on the fringe of a social movement contribute substantially to the
overall level of protest. We add to this discussion by demonstrating how
sociolinguistic cohesion evolves in response to negative and positive exogenous
events.

### Cohesion in Social Movements and Civil Wars
People join social and political resistance movements because they feel a sense
of connection to the cause, and because they have been sufficiently convinced to
participate. Given that people are generally risk averse, overcoming inertia
requires a strong connection to the group identity, or against an individual,
event, or political injustice [@Goldstone_1980; Goldstone_2011]. Political
resistance offers participants the opportunity to “lose themselves” and
relinquish their individual identity in favor of identification with a larger
group identity, thus contributing to a cohesive group ready and mobilized for
action [@Hoffer_2011. Previous research on social movements supports the idea
that participants are likely to join movements when they know and have a
personal connection to someone else already involved [@McAdam_1986. The type of
collective contentious behavior that manifests in society is a function of the
available political opportunity structures. The political opportunity structure,
as defined by Kitschelt (1986) are “comprised of specific configuration of
resources, institutional arrangements and historical precedents for social
mobilization, which facilitate the development of protest movements in some
instances and constrain them in others [@Kitschelt_1986].” What emerges as
strategies and tactics in contentious politics is in essence, a function of what
the system makes possible. This includes the number of people and their
disposition, i.e. nonviolent or violent, as well as their internal structure.
The more cohesive a group is, the greater threat they pose to the establishment.

Cohesion is also a familiar concept in the study of civil conflicts as it
pertains to group affiliation and the ability to overcome collective action
problems. Olson’s (Olson, 1965) seminal work on collective action informs the
theory of group behavior. Individuals must act in concert – cohesively – to
accomplish the task of challenging the authority of the state. Building on
Eisinger’s (1973) analysis of social protest movements [@Eisinger_1973, McAdam,
Tilly and Tarrow [@Tilly_2004, and Francisco [@Francisco_1995 have provided
foundational work on the dynamics of social movements and contentious behavior.
More recently, Gates (2002) considers rebel recruitment for participation in
civil conflicts as a function of internal cohesion [@Gates_2002]. He notes that
rebel movement success is influenced by the degree of group cohesion, which
itself is a function of geographic distance, ideology and ethnicity.

There are multiple rebel groups within Syria, some with fighters sourced locally
from the disaffected Syrian population, and an increasing number of foreign
fighters from transnational terrorist organizations. Gates’ work helps us
understand the intergroup competition for participation in the rebel movement,
as well as the groups’ abilities to punish defectors for nonparticipation.
Foreign fighters would appear to be amply motivated to fight, given their
decision to relocate from their home country which is sometimes in the region
but other times requires them to travel large distances.

Similarly, Sambanis and Shayo (2013) find in their model that “intense ethnic
conflict makes people care more about their ethnic group relative to other
groups, and seek to resemble it more [@Sambanis_Shayo_2013].” As ethnic conflict
intensifies, groups seeking to become more quintessentially like the typical
member of their ethnic group, this should be manifested in their language on
social media. Social medial should reflect this delineation amongst ethnic
groups and between them. However, this may also include a general identification
across ethnic groups in a membership of an opposition to an oppressive leader.

It is this general identification with the larger opposition movement, not the
individual identity groups within Syria, which we seek to understand in this
paper. It is implausible to expect there to be one cohesive narrative from the
multiple groups with disparate incentives. However, they do have important
similarities, namely, they are all interested in opposing the current Syrian
regime using largely violent means, although there are notable and regular
instances of nonviolent protests on Fridays, following prayer services. Thus,
what we are able to capture in this study is the degree to which all opposition
movements to the Syrian government have a similar content and tone, with a
similar ebb and flow. It should be increasingly threatening to the regime if the
rebel movement across the disparate groups is growing stronger and more unified,
represented by similar language used across groups.

The civil war literature has further explored the role of social cohesion in
detail. Collier and Hoeffler identified the role of social homogeneity as having
an increased likelihood on social conflict, especially when the society is
polarized [@Collier_Hoeffler_2004]. Cederman and Girardin  write that, “In
general, nationalist groups derive their cohesion from collective-level
identities that resonate deeply with the emotional needs of modern citizens.”
Sambanis and Schulhofer-Wohl note that “group cohesion and solidarity (underlie)
the offensive advantage (Sambanis & Schulhofer-Wohl, 2009).” Anyanwu (2002)
writes, “Another source of rebel military opportunity is social cohesion…an
ethnically diverse society may reduce the opportunity for rebellion by limiting
the recruitment pool [@Anyanwu_2002].” Referencing Gurr’s work on minorities at
risk, Regan and Norton (2005) note that “a religious or linguistic minority
might suffer disproportionately in a given society, and this form of
aggrievement can lead to unrest across the social lines that distinguish the
minority group. Therefore, while collective cohesion is a function of targeted
maldistributions, it is individual penury that allows for this collective
mobilization [@Gurr_2000; @Regan_Norton_2005].” Thus scholarship concurs that
group cohesion matters to group mobilization; we add to this concept by
introducing the role of language in facilitating cohesion.

## Sociolinguistic cohesion: Connectedness of speakers, events, and ideas
We tie the constructs of lexical, interlocutor, and social/physical cohesion
together in a theoretical framework called sociolinguistic cohesion, as shown in
Figure X. We can measure the changes in cohesion preceding and in response to
external stimuli by evaluating the messages and the context in which they occur.
Several mechanisms are in motion here. Previous values of linguistic cohesion
should influence future values of linguistic cohesion. They should also
influence future actions in the real world space. Previous events influence
future events, and also the future cohesiveness of the opposition.

Sociolinguistic cohesion is not a static concept. Just as social movements
experience increased participation as well as setbacks – indicating dynamic
relationships between participants, and participants and the sociopolitical
environment – sociolinguistic cohesion is also dynamic, reflecting the changing
states of the political sphere. Other scholars have noted the diffusion of
messages whereby they reach a critical mass, whereby “behaviors, beliefs, and
preferences” spread within a population [@Barash_Cameron_Macy_2012].  Cohesion
changes over time in response to external events and the internal dynamics of
the social movement. Some may wonder whether social cohesion can diffuse among
distant actors who do not interact face-to-face; Centola and Macy show that
Internet weak ties extend over long distances and as such have spatially
significant abilities to influence actors [@Centola_Macy_2007]. Fink et al. also
explored diffusion of hashtags in viral contagion in Nigeria, a phenomenon
similar to what we identify as sociolinguistic cohesion
[@Fink_Schmidt_Barash_Cameron_Macy_2016]. The prevalence and rate of particular
hashtags contribute to the overall sociolinguistic cohesion. he more that people
reuse hashtags or common phrases, the greater the sociolinguistic cohesion when
this phenomenon occurs during socially significant times like protests
[@Adamic_others_2015]. Exogenous events, like intervention by another country or
international organization like the United Nations, can affect sociolinguistic
cohesion. Endogenous events, like the recruitment of new members to the rebel or
resistance movement, can also affect the level of cohesion.

Combining these separate threads, it is reasonable to expect that real-world
actions and online social interactions would be enmeshed, driving and being
driven by one another. Rather than expecting that only real-world social
cohesion would be driven by real-world action, we would expect interaction
between real-world and virtual dynamics: In this case, exposure to real-world
violence could result in social cohesion in the virtual world. Increased social
cohesion is thus potentially an indicator and a result of increased social
organization and mobilization on the ground.

Exploring how social media influences contentious behavior, Tilly notes that
social media lowers coordination and transaction costs for participants, and
that older versions of organizing social movements have become less effective,
yielding to the new frontier of online organizing [@McAdam_Tarrow_Tilly_2003].
Koopmans extends this further, writing that, “Authorities react to social
movement activities if and as they are depicted in the mass media, and
conversely movement activists become aware of political opportunities and
constraints through the reactions (or non-reactions) that their actions provoke
in the public sphere” [@Koopmans_2004].

Lexical cohesion, then, should be a useful indicator of social cohesion in
political crisis. Lexical cohesion happens at the word level. Interlocutor
cohesion happens between speakers. Political cohesion happens at the social
level during moments of political crisis, like civil wars and social movements.
Sociolinguistic cohesion is comprised of lexical, interlocutor, and
socio-political events, where the words and their context reveal the disposition
of socially connected linguistic communities. Sociolinguistic cohesion combines
the elements of lexical cohesion in that it captures the degree of similarity
between what is said, and how it is said. It accounts for the cohesion between
multiple speakers in an environment whose language evolves to become more
aligned and similar. Sociolinguistic cohesion also addresses how groups involved
in political rebellion must have internal cohesion to overcome the collective
action problem to mount sustained resistance to state authority.

Social coordination has long been used both physically and virtually to start
and carry out collective actions. These collective events bring individuals
together and create emergent patterns of communication and action. Online
communication has historically been used to document and communicate archival
information about events, but it has been increasingly used to bring about
social change. From a dynamical systems perspective, these different means of
social coordination during times of turmoil can be considered two complementary
components of a single system, mutually constrained and constantly interacting
and influencing one another [@mahner1997foundations]. Investigation of the
coupling of real-world mobilization and social cohesion could provide new
insight into the power of social media during mobilization and the ways in which
individuals utilize it to bring about, or respond to, social change.
Increasingly, social media platforms are being utilized during social
mobilization, but the relationship between virtual and real-world mobilization
is still largely unknown. The present work aims to understand how physical and
virtual movements coexist and interact from a dynamical systems perspective.

Just as real-world social cohesion can be quantified, so can online cohesion.
While this can be quantified in many ways, one kind of cohesion can be found in
the similarity in language use on social media. This implicit or explicit
similarity in language use can indicate shared beliefs, thoughts, and feelings,
as has been studied in research on language style in psycholinguistics and
social psychology [@gonzales2010language; @tausczik2010psychological;
@brennan1996conceptual]; although previous work on language style has often
focused on individuals and small groups, this can be scaled up to examine
massive social-level changes [@michel2011quantitative; @fusaroli2015timescales].
This kind of widespread online social cohesion can lead to real-world social
change.

This kind of massive online social cohesion can lead to real-world social
manifestations. Social media platforms provide one way to enable this kind of
online social cohesion by offering distributed information-gathering and
real-time information dissemination. Given its accessibility and proficiency as
an information exchange tool, Twitter has provided individuals involved in
social movements an opportunity to complement or even replace real-world social
mobilization. For example, during the 2012 Italian protests for global economic
change, the riots in Rome were mirrored by Twitter activity both on the day of
the protests and in the days following them [@vicari2013public]. Tweeting was
determined to be more effective locally in discussing real-world, real-time
events than traditional mainstream media platforms. As more and more individuals
rely on social media to obtain local news and information, it becomes
increasingly important to understand just how these micro-blogs reflect
real-world action.

## Using Nonlinear Methods to Capture Complex Relationships
Given the hypothesized interrelationship between the real-world and online
actions---and the notorious messiness of naturally occurring data---it is
critical to find appropriate analysis tools that can identify the complex
interconnections. Traditional linear analyses have led to many notable
discoveries in the domains of psychology and political science, but time series
data pose problems that are not easily overcome with these methods. Events have
historically been conceptualized as linear chains of smaller, sequential events
with one occurrence leading to the next. Certain kinds of data are relatively
well-suited to these analyses, but messy, complex, variable data from real-world
events can pose challenges, from overpowered samples to violation of underlying
statistical assumptions [@paxton2017finding].

At the same time, the foundational theoretical assumptions of a linear world has
been increasingly replaced by ideas from dynamical systems theory. Rather than a
chain of events, time series data can be thought of as nested occurrences, where
mutually constrained parts (e.g., citizens, governments, real-world events,
social media platforms) are constantly interacting, leading to more global
behaviors that are larger than the sum of the parts [@turvey2018lectures].
Linear analyses flatten time series data and ignore the variability that is
fundamental to these systems and their evolution over time. However, when
analyzed as dynamical systems, many phenomena exhibit new nonlinear patterns
that were formerly overlooked. Nonlinear methods provide means for capturing the
rich temporal dynamics and variability that unfold in time series data.

Although a number of other nonlinear methods exist, the primary nonlinear method
used in the current work is cross-recurrence quantification analysis [CRQA;
@zbilut1998detecting] and its offshoot, windowed cross-recurrence quantification
analysis [WCRQA; @webber2005recurrence]. CRQA has successfully been used to gain
a better understanding of many complex human phenomena, including a variety of
human social interactions [for a review, see @fusaroli2014analyzing]. Much as it
has for psychology, we argue that the use of nonlinear methods in the
investigation of political science phenomena will lead to new insights and the
development of more robust theories.

## The Present Work
Social media has become a primary platform for social movements and activism
across the world, but the extent to which virtual activism reflects and drives
real-world mobilization is still largely unknown. In the current work, we
conceptualize the physical and virtual movements as a dynamical system, both
constantly interacting and influencing the other. We aim to study this by
analyzing the real-world events and online social cohesion during nearly three
months of the Arab Spring. Both real- and virtual-world collective coordination
were powerful tools in the events of the Arab Spring and are undeniably linked.
This study aims to explore the dynamics of Syrian Twitter users, who used the
social media platform to raise awareness about the Arab Spring and bring about
mass mobilization.

Using cross-recurrence quantification analysis [CRQA; @zbilut1998detecting] and
windowed cross-recurrence quantification analysis [WCRQA;
@webber2005recurrence], we analyzed the co-evolution of social cohesion online
and real-world events. Specifically, using heterogeneous data from March 30,
2012, to June 15, 2012, we compared the social cohesion of Syrian tweets with
the counts of all, positive, and negative real-world events derived from
gold-standard international event data [i.e., the Integrated Crisis Early
Warning System; @DVN/28075_2015].

We hypothesized that the count of all events and the count of negative events
would show high levels of determinism, as this time period was marked by daily
violence which often leads to increased salience. We also hypothesized that the
count of positive events would not show patterns of determinism, as these events
were likely overshadowed by the salient negative daily occurrences. After
testing these hypotheses, we then conducted exploratory analyses to further
examine the patterns identified with CRQA. Using WCRQA, we sought to identify
shifts in the dynamics of the systems that could then be linked to real-world
shifts in conflict frequency. Through these analyses, our two primary goals in
the current work were to uncover (1) how different intensities during times of
turmoil may enhance or deter social cohesion in social media communication and
(2) how social mobilization may be evident in the dynamic coupling of cohesion
and real-world events.

An additional focus of the current work is to introduce cross-recurrence
quantification analysis---a longstanding method from physics that has since
become influential in a variety of other fields---to political science. We do so
by providing detailed descriptions of the methods and providing clear directions
about interpretation and significance testing even with relatively small-*n* or
case studies. After demonstrating these methods' utility with this dataset, we
close the paper with specific suggestions to political scientists interested in
incorporating these methods into their work.

# Method
## Materials
### Tweet Corpus

```{r tweet-count, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, results='hide'}

# read in social cohesion data
cohesion_df <- read.csv("./data/raw/Syria-social_cohesion.csv")

```

The corpus of tweets used included `r sum(cohesion_df$n)` Syrian tweets
occurring between March 30--June 15, 2012. Using the Twitter streaming API,
servers at the University of Texas at Austin collected the tweets. The servers
were run on a stable (LTS) version of Ubuntu Linus (v. 10.04) and were connected
directly to the high-bandwidth UT network maintained by IT services.

Tweets collected contained one or more of the following keywords: “syria,”
“syrian,” “damascus,” “homs,” “al-assad,” and “sunni.” These keywords were
chosen based on the most distinctive terms used in news articles covering the
Arab Spring in Syria at the time. A node.js script was used to reinitialize the
connection to Twitter's servers whenever the script timed out due to a period of
silence. The script requested data from the API endpoint using basic HTTP
authentication (OAuth was optional, at the time) at
https://stream.twitter.com/1.1/statuses/filter.json, and stored the data
directly to the local hard disk. The tweets were then post-processed with a
Python script to convert the JSON-formatted tweets to tab-separated files, which
were then compressed to maximize storage efficiency.

All scripts used for scraping tweets and compressing data can be found at
https://github.com/chbrown/twilight.

### Event Data
Event data were obtained from the 2012 Integrated Crisis Early Warning System
[ICEWS; @DVN/28075_2015]. The event data were first filtered to remove
incomplete and incorrectly formatted data. They were then filtered to include
only events that were directed at (i.e., `target`) Syria ($n = 6300$). To
explore the impact of outwardly directed Syrian events, the data were filtered
again to include events that were both directed at (i.e., `target`) or directed
by (i.e., `source`) Syria ($n = 7990$).

## Data Preparation
### Social Cohesion Metric
The social cohesion or linguistic similarity of daily tweets was calculated by
quantifying the frequency of shared content words with respect to those words’
frequencies in the entire tweet corpus. If a linguistic term has a high
frequency in one tweet but has a low frequency in the whole tweet corpus, then
it is more likely a content word and is thus given more weight in calculating
social cohesion. This method is similar to inverse document frequency (IDF). For
more information on IDF, see @jones1972statistical and
@robertson2004understanding.

The tweets were first sorted into ascending order by timestamp. They were then
grouped into successive sets of five tweets. Inverse frequency weighting for the
overlap words and all words was conducted for each pair of two tweets in the
group of five, then divided to get the social cohesion score for that pair (see
Eq. 1-3). This resulted in ten social cohesion values for each group of five
tweets; these values were then averaged to get a social cohesion score for the
entire set of five tweets. Finally, the mean of all average cohesion values for
all groups was taken to get a social cohesion score for each day.

\begin{align}
\text{Inverse frequency weighting of overlap words = }\sum_{i=1}^k
\frac{log(f(t_i, d1) + f(t_i, d2) + 1)}{log(f(t_i) + 1)}
\end{align}

\begin{align}
\text{Inverse frequency weighting of all words = }\sum_{i=1}^n \frac{log(f(t_i,
d1) + f(t_i, d2) + 1)}{log(f(t_i) + 1)}
\end{align}

\begin{align}
\text{Two tweets' similarity =}\frac{\text{inverse frequency weighting of overlap
words}}{\text{inverse frequency weighting of all words}}
\end{align}

The social cohesion calculation was performed on the tweet corpus containing
both English and Arabic tweets. All hashtags and URLs were included in the
analysis.

### ICEWS Event Time Series
Using the Integrated Crisis Early Warning System (ICEWS) developed at the
Defense Advanced Research Project Agency, the daily events were scored with an
intensity rating ranging from -10 to 10. Negative events---such as riots and
attacks---are coded with low negative numbers, while positive or peaceful events
are coded with high positive numbers. This well-validated coding scheme allows
for numerical quantification of different event types.

For both the target-only and the target and source data described above, time
series of the count of all events were generated by summing the total number of
events on each day, irrespective of ICEWS score. Similarly, time series of count
of positive events were generated by summing the number of ICEWS events for each
day that had an intensity rating greater than zero. Lastly, time series of count
of negative events were generated by summing the number of ICEWS events for each
day that had an intensity rating less than zero.

## Analyses
### Cross-Recurrence Quantification Analysis
We first used cross-recurrence quantification analysis [CRQA;
@zbilut1998detecting] to investigate the temporal patterns of social cohesion
and real-world events in Syria during Arab Spring. CRQA is an extension of RQA,
which captures the patterns of revisited states of a single time series
[@zbilut1992embeddings; @webber1994dynamical]. CRQA allows for two time series
to be projected onto one another to investigate shared states and trajectories.
In CRQA, we track these shared states by identifying *recurrent points*, or
intersections of time during which two signals shared or revisited the same
state. These shared states can be at the same time or at later points in the
time series. We can further chart shared trajectories by examining the
occurrences and patterns of multiple sequential recurrent points.

To perform CRQA, the time series must be converted to the same scale, while
still preserving a sampling rate that respects the native timescale of the
system [cf. Nyquist frequency in signal processing; @grenander1959probability].
Because the social cohesion and event data have different scales and values, we
converted each time series (e.g., Figure \@ref(fig:raw-ts)) into deciles (e.g.,
Figure \@ref(fig:deciled-ts)) to analyze the relative dynamics of different
levels of social cohesion and counts of all, positive, and negative events on
each day. We implemented CRQA using the `crqa` package [@R-crqa] for R
[@R-base].

#### Visualization
A cross-recurrence plot (CRP) is the visual representation of CRQA. To construct
a cross-recurrence plot, two time series are plotted against each other on the
$x$ and $y$ axes. At every point where they share the same state, a point in
placed on the plot to indicate recurrence. (For categorical CRQA, this process
is similar to the Cartesian product.)

We can visually analyze these plots to reveal patterns of structure and
recurrence. The line $y = x$ is known as the *line of identity*. This line falls
where the time series match temporally. A filled line would indicate perfect
in-phase synchrony between the two time series. See Figures
\@ref(fig:plot-rp-targ-all)-\@ref(fig:plot-rp-source-targ-neg) for examples of
RPs.

#### Metrics
While visual inspections can be informative, we can derive a variety of
quantitative metrics from the cross-recurrence plot. Each of these metrics
provides information into the patterns and dynamics of the new shared system.
These metrics are also critical for inferential statistics. We will focus most
directly on two of these metrics: *recurrence rate* and *determinism*.

The percentage of recurrent points to total possible points is a metric called
the *recurrence rate* (*RR*). This metrics gives a general description of how
frequently the time series are sharing the same states. In the current study, a
recurrent point would indicate that the level of social cohesion and frequency
of events were at the same intensity.

Recurrent points that occur in successive moments form line structures. These
are conceptualized as shared trajectories, where the two time series are moving
together across time in a shared state. The percentage of recurrent points on
the plot that fall along diagonal line structures (i.e., two or more consecutive
points) is a metric called *percent determinism* (or simply *determinism*;
*DET*). For the current project, a high DET would indicate that not only are the
levels of social cohesion and frequency of events occurring at similar
intensities, but they are doing so across consecutive days.

A number of other metrics provide insight not only into the amount of structure
between the systems but about the nature of this shared structure. The average
length of the line structures is called *mean line* (L), which can be
conceptualized as the average amount of time that the two time series move
together before one or both shift states. The longest line length is called *max
line* (*maxL*); it indicates the longest instance of being “stuck” in a
particular state together and is a measure of the stability of the system. The
total *number of lines* on the plot (*NRLINE*) scales with structure: The more
lines that exist, the more coupled the two systems are. The complexity or
variability of line lengths is captured by a metric called *entropy*
(*ENTR*)---which may also be normalized by the total number of lines (i.e.,
*normalized entropy*; *rENTR*)---and indicates the stability and predictability
of the system. When one of the signals gets “stuck” in a state and the other
moves on to other states, this results in vertical line structures, which are
quantified in *laminarity* (*LAM*; i.e., the proportion of points forming
vertical line structures) and *trapping time* (*TT*; i.e., the average length of
the vertical trajectories). For more detail on CRQA metrics, see @R-crqa.

#### Parameters
To conduct the analyses on the decile data, the parameters were set to those
typically used in categorical CRQA: Delay was set to $0$, embedding dimension
was set to $1$, and radius was set to $.001$ to allow for only exact recurrent
matches. No normalization was done, and lines were considered two or more
consecutive recurrent points. The Theiler window was set to $0$ to include
points along the line of identity. For more details on the CRQA parameters and
how they impact the analysis, see @R-crqa.

#### Permutation Test
Unlike some nonlinear analyses (e.g., fractal analyses), CRQA metrics do not
have inherently meaningful values. As a result, CRQA is best used as a relative
metric---that is, comparing the CRQA values across conditions or subsets of the
data. Typically, meaningful differences between values are determined with
traditional inferential statistics. While this may not be a problem for
experimentally derived datasets, it can pose a barrier to researchers using
small-$n$ data or conducting data-informed case studies.

For CRQA, this can be addressed using *permutation tests*. Similar to more
established methods of statistical significance in CRQA research [which create
surrogate time series; e.g., @paxton2017interpersonal], the permutation test
approach to significance testing allows the researcher to use a time series as
its own baseline by testing whether the *structure* of the unfolding events of
the time series---rather than the *raw frequency* of the events within the time
series---cohere together more than would be expected by chance. A permutation
test is similar to bootstrapping, except that the sampling of observations is
done without replacement. In this approach, we simply create a large number of
permutations of the original time series (i.e., breaking up the dependency
across time points but preserving the frequencies of the events), conduct CRQA
for each of those permutations, and then examine whether the real time series'
CRQA metrics exceed those that we would expect to see by chance. The proportion
of times that the real time series' values exceeds the surrogate time series'
values can be interpreted as the alpha criterion to determine significance. For
more information on permutation tests, see @good2013permutation.

Using the real deciles of social cohesion, count of all events, count of
positive events, and count of negative events, we conducted a series of
permutation tests (one for the comparison between social cohesion and each type
of event count) to determine statistical difference from chance for each CRQA
metric. Using the `sample` function in base R [@R-base], $1000$ permuted time
series were created for each of the real time series. CRQA was then run $1000$
times for social cohesion and count of all events, social cohesion and count of
positive events, and social cohesion and count of negative events using the
permuted time series. This resulted in $1000$ results for each of the CRQA
metrics for each combination of time series. Statistical significance from
chance was set at the 95th and 99th percentiles of the metrics from the permuted
CRQ analyses. Given that a permutation test is similar to bootstrapping but does
not include replacement, all points were included in each CRQA analysis, leaving
RR constant across all permutations. Thus, RR is not included in the inferential
tests.

### Windowed Cross-Recurrence Quantification Analysis
We used windowed cross-recurrence quantification analysis [WCRQA;
@webber2005recurrence] to further investigate the evolution of the temporal
dynamics of social cohesion and real-world events. In this analysis, a smaller
“window” of time points is used to calculate CRQA and capture more fine-grained
dynamics of two time series. The window remains the same size and “slides”
diagonally up the line of identity. The changes in the CRQA metrics across the
windows can reveal changes in dynamics that can then be linked to external,
real-world events. We implemented WCRQA using the `crqa` package [@R-crqa] for R
[@R-base].

#### Metrics
For WCRQA analyses, we decided to focus on RR and DET.

#### Parameters
In addition to the CRQA parameters, two other parameters are required when
conducting WCRQA. First, the *window size* will determine how many observations
are used in each CRQA analysis. We conducted WCRQA with a window size of $14$ to
investigate the dynamics over two weeks and allow for the capture of more
structure and thus a more meaningful interpretation of DET.

The second parameter to be chosen is a *window step size*. Window step size is
how far the window moves up the line of identity before conducting CRQA again.
We chose $1$ for the step size to understand how the metrics evolve from day to
day.

All other CRQA metrics remained the same as those previously noted.

#### Visualization
To visualize WCRQA, the resulting metrics from each window can be plotted as a
time series. This allows for visual interpretation of the pattern or fluctuation
of the metrics across time. These dynamics can then be linked to other external
events. See Figures \@ref(fig:plot-wcrqa-targ-all) and
\@ref(fig:plot-wcrqa-source-targ-all) for examples.

#### Permutation Test
To visualize the differences in the WCRQA metrics from those obtained by chance,
a permutation test was done to set upper and lower bounds for the 95th and 99th
percentiles. Using a sample size of 14, $1000$ permuted time series were
generated without replacement for each real time series. The upper and lower
95th and 99th percentiles were set based on the distribution of the permuted
WCRQA metrics.

# Results
```{r primary-crqa, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, results='hide'}

# run primary crqa analysis
source('./scripts/primary-analyses/02_crqa.R')

# read in metric names
metrics <- t(read.csv("./results/primary/crqa/target-crqa_results-all_events.csv", 
                      header = FALSE)[-2,-1])

# read in metrics from crq analyses (target filtered)
metrics_all_targ <- t(read.csv("./results/primary/crqa/target-crqa_results-all_events.csv", 
                               header = FALSE)[-1,-1])
metrics_pos_targ <- t(read.csv("./results/primary/crqa/target-crqa_results-pos_events.csv", 
                               header = FALSE)[-1,-1])
metrics_neg_targ <- t(read.csv("./results/primary/crqa/target-crqa_results-neg_events.csv", 
                               header = FALSE)[-1,-1])

# read in significance from permutation tests (target filtered)
sig_all_targ <- t(read.csv("./results/primary/crqa/target-metric_signif_all.csv", 
                           header = FALSE)[-1,-1])
sig_pos_targ <- t(read.csv("./results/primary/crqa/target-metric_signif_pos.csv", 
                           header = FALSE)[-1,-1])
sig_neg_targ <- t(read.csv("./results/primary/crqa/target-metric_signif_neg.csv", 
                           header = FALSE)[-1,-1])

# read in metrics from crq analyses (source and target filtered)
metrics_all_source_targ <- t(read.csv("./results/primary/crqa/source_target-crqa_results-all_events.csv",
                                      header = FALSE)[-1,-1])
metrics_pos_source_targ <- t(read.csv("./results/primary/crqa/source_target-crqa_results-pos_events.csv",
                                      header = FALSE)[-1,-1])
metrics_neg_source_targ <- t(read.csv("./results/primary/crqa/source_target-crqa_results-neg_events.csv",
                                      header = FALSE)[-1,-1])

# read in significance from permutation tests (source and target filtered)
sig_all_source_targ <- t(read.csv("./results/primary/crqa/source_target-metric_signif_all.csv", 
                                  header = FALSE)[-1,-1])
sig_pos_source_targ <- t(read.csv("./results/primary/crqa/source_target-metric_signif_pos.csv", 
                                  header = FALSE)[-1,-1])
sig_neg_source_targ <- t(read.csv("./results/primary/crqa/source_target-metric_signif_neg.csv", 
                                  header = FALSE)[-1,-1])

```

```{r raw-ts, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = 'Example of pre-processed data.'}

# read in processed data
data <- read.csv("./data/formatted/primary/formatted_data.csv", header = TRUE)

raw_ts_example <- ggplot(data, aes(x= as.Date(Date), y=all_events_target)) +
  geom_point() +
  geom_line() +
  xlab("Date") +
  ylab("Count of events")
ggsave("./results/primary/tally_ts_example.png")

# call the image
knitr::include_graphics("./results/primary/tally_ts_example.png")

```

```{r deciled-ts, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = 'Example of deciled data.'}

# plot the deciles time series and save
deciled_ts_example <- ggplot(data, aes(x= as.Date(Date), y=all_deciles_target)) +
  geom_point() +
  geom_line() +
  xlab("Date") +
  ylab("Count of events (deciles)")
ggsave("./results/primary/deciled_ts_example.png")

# call the image
knitr::include_graphics("./results/primary/deciled_ts_example.png")

```

## Cross-Recurrence Quantification Analysis

### Syria as Target

A permutation test for the CRQA of social cohesion and count of all events found
a statistical difference from chance for DET, NRLINE, maxL, and LAM. All other
metrics did not reach significance (Figure \@ref(fig:plot-rp-targ-all);
\autoref{table-1}).

A permutation test for the CRQA of social cohesion and count of positive events
found no statistically significant difference for any of the metrics other than
TT (Figure \@ref(fig:plot-rp-targ-pos); \autoref{table-1}).

A permutation test for the CRQA of social cohesion and count of negative events
found a statistically significant difference for DET, NRLINE, L, ENTR, LAM, and
TT. All other metrics did not reach significance (Figure
\@ref(fig:plot-rp-targ-neg); \autoref{table-1}).

A qualitative investigation of the recurrence plots suggests a shift in dynamics
around the 36th observation, which corresponds to May 4th, 2012.

```{r primary-crqa-table-target, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}

# bind metrics and significance results
target_results <- as.data.frame(cbind(metrics,
                                      metrics_all_targ,
                                      sig_all_targ,
                                      metrics_pos_targ,
                                      sig_pos_targ,
                                      metrics_neg_targ,
                                      sig_neg_targ), stringsAsFactors = FALSE)

# covert to numerics for rounding
numerics <- c(2:7)
target_results[,numerics] <- sapply(target_results[,numerics],as.numeric)

# add significance stars
stars_all <- as.data.frame(stars.pval(target_results[,3]))
stars_pos <- as.data.frame(stars.pval(target_results[,5]))
stars_neg <- as.data.frame(stars.pval(target_results[,7]))

target_results <- cbind(target_results, stars_all, stars_pos, stars_neg)
target_results <- target_results[,c(1,2,3,8,4,5,9,6,7,10)]

# give intuitive names
colnames(target_results) <- c("metric", 
                              "all events", "p", "sig.", 
                              "positive events", "p", "sig.", 
                              "negative events", "p", "sig.")

# exclude row names
rownames(target_results) <- NULL

# set to not split table and round to two decimal places
panderOptions("table.split.table", Inf)
panderOptions('round', 3)
panderOptions('keep.trailing.zeros', TRUE)

# style for Rmd
panderOptions("table.style", "rmarkdown")

# build table
pander(target_results, caption = "\\label{table-1}CRQA results for target-only data.")

```

```{r plot-rp-targ-all, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = 'Recurrence plot (RP) for social cohesion and count of all events from target filtered data.'}

# call image
knitr::include_graphics("./results/primary/crqa/target-rp-all_events.png")

```

```{r plot-rp-targ-pos, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = 'Recurrence plot (RP) for social cohesion and count of positive events from target filtered data.'}

# call image
knitr::include_graphics("./results/primary/crqa/target-rp-pos_events.png")

```

```{r plot-rp-targ-neg, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = 'Recurrence plot (RP) for social cohesion and count of negative events from target filtered data.'}

# call image
knitr::include_graphics("./results/primary/crqa/target-rp-neg_events.png")

```

### Syria as Source and Target
A permutation test for the CRQA of social cohesion and count of all events found
a statistical difference from chance for DET, NRLINE, and LAM. All other metrics
were non-significant (Figure \@ref(fig:plot-rp-source-targ-all);
\autoref{table-2}).

A permutation test for the CRQA of social cohesion and count of positive events
found a statistical difference from chance for maxL, ENTR, and LAM (Figure
\@ref(fig:plot-rp-source-targ-pos); \autoref{table-2}).

A permutation test for the CRQA of social cohesion and count of negative events
found a statistical difference from chance for DET, NRLINE, and rENTR. All other
metrics were non-significant (Figure \@ref(fig:plot-rp-source-targ-neg);
\autoref{table-2}).

Looking qualitatively at the recurrence plots, there again appeared to be a
shift in dynamics around the 36th observation (May 4th, 2012).

```{r primary-table-source-target, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE}

# bind metrics and significance results
source_target_results <- as.data.frame(cbind(metrics,
                                      metrics_all_source_targ,
                                      sig_all_source_targ,
                                      metrics_pos_source_targ,
                                      sig_pos_source_targ,
                                      metrics_neg_source_targ,
                                      sig_neg_source_targ), stringsAsFactors = FALSE)

# covert to numerics for rounding
numerics <- c(2:7)
source_target_results[,numerics] <- sapply(source_target_results[,numerics],as.numeric)

# add significance stars
stars_all <- as.data.frame(stars.pval(source_target_results[,3]))
stars_pos <- as.data.frame(stars.pval(source_target_results[,5]))
stars_neg <- as.data.frame(stars.pval(source_target_results[,7]))

source_target_results <- cbind(source_target_results, stars_all, stars_pos, stars_neg)
source_target_results <- source_target_results[,c(1,2,3,8,4,5,9,6,7,10)]

# give intuitive names
colnames(source_target_results) <- c("metric", 
                                     "all events", "p", "sig.", 
                                     "positive events", "p", "sig.", 
                                     "negative events", "p", "sig.")

# exclude row names
rownames(source_target_results) <- NULL

# set to not split table and round to two decimal places
panderOptions("table.split.table", Inf)
panderOptions('round', 3)
panderOptions('keep.trailing.zeros', TRUE)

# style for Rmd
panderOptions("table.style", "rmarkdown")

# build table
pander(source_target_results, caption = "\\label{table-2}CRQA results for source and target data.")

```

```{r plot-rp-source-targ-all, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = 'Recurrence plot (RP) for social cohesion and count of all events from source and target filtered data.'}

# call image
knitr::include_graphics("./results/primary/crqa/source_target-rp-all_events.png")

```

```{r plot-rp-source-targ-pos, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = 'Recurrence plot (RP) for social cohesion and count of positive events from source and target filtered data.'}

# call image
knitr::include_graphics("./results/primary/crqa/source_target-rp-pos_events.png")

```

```{r plot-rp-source-targ-neg, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = 'Recurrence plot (RP) for social cohesion and count of negative events from source and target filtered data.'}

# call image
knitr::include_graphics("./results/primary/crqa/source_target-rp-neg_events.png")

```

## Windowed Cross-Recurrence Quantification Analysis
```{r primary-wcrqa, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, results='hide', include = FALSE}

# run primary crq analysis
source('./scripts/primary-analyses/03_windowed_crqa.R')

```

In the plots (see Figures \@ref(fig:plot-wcrqa-targ-all) and
\@ref(fig:plot-wcrqa-source-targ-all)), the 95th percentiles are represented by
horizontal red lines, and the 99th percentiles are represented by horizontal
orange lines. The trend lines are plotted in blue.

### Syria as Target

```{r plot-wcrqa-targ-all, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = 'Windowed CRQA for social cohesion and event counts from target filtered data.', out.extra='angle=90'}

# call image
knitr::include_graphics("./results/primary/windowed-crqa/target-windowed_all.png")

```

```{r plot-mode-event, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = 'Mode ICEWS event per day.'}

mode_plot <- ggplot(data, aes(x= as.Date(Date), y=mode_source_target)) +
                    geom_point() +
                    geom_line() +
                    xlab("Date") +
                    ylab("Mode ICEWS event")
ggsave("./results/primary/mode_ts_plot.png")

# call the image
knitr::include_graphics("./results/primary/mode_ts_plot.png")

```

For social cohesion and count of all events, there was a consistent period of
above-chance RR from the 7th--12th and 36th--47th window. In addition to some
one-week peaks at the beginning of the time period, DET was largely above chance
from the 46th to the 61st window, with some exceptions (Figure
\@ref(fig:plot-wcrqa-targ-all)).

For social cohesion and count of positive events, we saw a downward trend in RR
as the window advances throughout the sample. Interestingly, we see largely
above-chance RR until the 50th window, with some exceptions during window
periods 20-30. DET was at or above chance for the 7th--17th, 32nd--34th, and
60th windows (Figure \@ref(fig:plot-wcrqa-targ-all)).

For social cohesion and count of negative events, RR stayed above or almost
above chance from the 4th--7th and 43rd-47th windows. DET was above or nearly
above chance in the 1st and 2nd windows, with intermittent peaks in later
windows (Figure \@ref(fig:plot-wcrqa-targ-all)).

### Syria as Source and Target

```{r plot-wcrqa-source-targ-all, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, fig.cap = 'Windowed CRQA for social cohesion and event counts from source and target filtered data.', out.extra='angle=90'}

# call image
knitr::include_graphics("./results/primary/windowed-crqa/source_target-windowed_all.png")

```

For social cohesion and count of all events, there was a consistent period of
above- or nearly above-chance DET from the 7th--13th and 26th--47th windows. DET
also showed above- or nearly above-chance values in windows 7 through 17 and
intermittent peaks between the 34th and 61st windows (Figure
\@ref(fig:plot-wcrqa-source-targ-all)).

For social cohesion and count of positive events, there was a downward trend in
RR as the window moved, with largely above-chance values in the first 20 windows
and the 32nd--45th windows. DET stayed largely within chance, with the exception
of windows 13 through 19 (Figure \@ref(fig:plot-wcrqa-source-targ-all)).

For social cohesion and negative events, there was an extended period of time
during which RR (32nd--47th windows) and DET (35th--39th windows) were above or
nearly above chance (Figure \@ref(fig:plot-wcrqa-source-targ-all)). These
windows span from May 3rd--14th, 2012.

# Discussion
In the present study, we investigated the dynamics of real-world events and
online communication on Twitter in Syria during Arab Spring. CRQA methodology
allows us to model the relationship between sociolinguistic cohesion and
exogenous political events. This work stands alongside the model of
Palestinian-Israeli social media and events developed by Zeitzoff (2011), and
the role of peripheral participants in the Syrian Arab Spring movement by
Steinert-Threlkeld (2017). By introducing the CRQA and WCRQA methods, we
demonstrate how language and social mobilization interact with events along the
conflict-cooperation continuum in Syria. We have introduced a novel theoretical
framework of sociolinguistic cohesion which we operationalize using a weighted
measure of lexical relationships. We examined how social mobilization both
virtually and physically evolve over time in tandem. Using nonlinear dynamical
time series analyses, we investigated the theory that virtual mobilization is
increasingly reflective of real-world events during times of strife.

## Cross-Recurrence Quantification Analysis
### Syria as Target
The significant result of DET for social cohesion with all and negative events
supports our first hypothesis that event salience is coupled with online social
cohesion. The frequency of all and negative events are thus associated with
greater shared trajectories with social cohesion that would be expected by
chance. However, when we specifically target positive events, the connection to
social cohesion fades, providing further support for our second hypothesis.
Positive events may be less coupled to social cohesion than more salient
negative events during times of turmoil; by contrast, overall event frequency
and negative event frequency are coupled with Twitter social cohesion.

### Syria as Source and Target
By adding Syria as a source in the data filtering, the dataset increased in size
by 1,690 samples. These additional observations are times where Syria initiated
an event with another country. Events initiated by Syria but directed internally
within the country were already captured by the target-only filtering. The
increase in coupling between positive events and social cohesion can thus only
be due to the externally directed events. Though the coupling between social
cohesion with negative and all events were largely unchanged, positive events
and social cohesion showed greater stability in their coupling (reflected in
maxL); while TT was no longer significant in these analyses, LAM (related to TT)
did reach statistical significance.

While interesting, the underlying meaning or drivers of this change in results
requires further investigation that incorporates the actual content of the
tweets themselves. One possible reason (among many) may be that the finding was
spurious. The positive external events from Syria may have been self-serving,
aiming to prevent intervention by other countries or to cover up what was
happening internally. During these moments, the increased cohesion may reflect a
surge from the Syrian people trying to raise awareness of the situation or to
decry the Syrian government's attempts to improve their international standing.
In this case, the increased coupling of positive events would be coupled not
with shared positive affect in social media but instead to shared negative
affect in social media in response. However, we cannot test this hypothesis with
the current data; future research should explore the semantic content of this
tweet corpus to determine the nature of the socially cohered language.

## Windowed-Recurrence Quantification Analysis

### Syria as Target
The time period of differing dynamics noted in the cross-recurrence plots---from
observation 32 through 47---appears to be captured by the above-chance RR values
in WCRQA of social cohesion and count of all events (Figure
\@ref(fig:plot-wcrqa-targ-all)). However, DET for these periods only
intermittently rises above chance; it more consistently reaches statistical
significance only at the end of the window through the end of the observation
period, following a spike in the RR between negative event count and social
cohesion. During this period, the modal ICEWS event for each day (Figure
\@ref(fig:plot-mode-event)) was almost consistently -10, the most violent event
category. This supports our first hypothesis.

Surprisingly---and contrary to our second hypothesis---we saw consistent
coupling between positive event count and social cohesion. Although both had a
downward trajectory throughout the time period under consideration, RR was at or
above chance in nearly all windows, and DET showed much more consistent shared
trajectories than negative events. This may suggest that Syrian Twitter
community was reacting more cohesively to the positive events aimed at Syria
from the international community; this interpretation is supported by the
attenuation of the effect in the analyses that include Syria as both a source
and a target.

### Syria as Source and Target
WCRQA on the source-and-target-filtered data showed similar patterns of RR
between the count of all events and social cohesion (as observed in the
target-only data), though the values increased. This suggests that large numbers
on a given day match increased levels of social cohesion online. However, the
results for DET changed: While the majority of statistically significant DET
windows occurred toward the end of the observation period for target-only data
(creating a positive trendline), the source-and-target data showed a negative
trendline for DET.

The divergences in the all-event count data for these two datasets underscores
the differences between RR and DET. As noted earlier, DET captures
*trajectories*---in this case, similar to the impact of actions and social
cohesion carrying across time. The same number of points (RR) can fall in
different patterns (DET), providing important information about the structure of
the system in question.

WCRQA metrics for positive-only event counts with social cohesion showed similar
patterns in RR and DET as the target-only data, but the precise values of DET
and RR were diminished. This may support the ideas suggested in the CRQA
section, but again, only future work including the *content* of those tweets can
directly speak to that explanation.

Finally, the coupling between negative-only event counts and social cohesion
also remained relatively similar but demonstrated some important differences.
First, we see a greater number of windows at above-chance levels for RR,
including an interesting spike around window 32 that suggests a negative
externally focused action taken by Syria prompted a sharp rise in social
cohesion. (This would not have been captured by the target-only dataset.)
Second---like with the all-event data---we see different structuring of the data
through differences in the patterns of DET; above-chance DET only occurrs within
the middle of the observation period, around a time of increased strive (as
noted above).

## Implications for Methodology and Theory

Nonlinear methods have been transformative for a variety of research areas
[including psychology, @coco2014cross]. One of the goals in the present work was
to introduce variants of one kind of nonlinear method---specifically, recurrence
quantification analysis---to this research area. While a number of excellent
guides can provide detailed explanations of the methods and their statistical
underpinnings [e.g., @riley2005tutorials; @webber1994dynamical;
@webber2005recurrence], one of the most important considerations that should be
briefly summarized here is about data.

Nonlinear methods---including recurrence-based analyses---require time series
data. This requires repeated observations of the behavior at interest,
preferably at a frequency that provides information about timing at a temporal
resolution that allows the researcher to make observations about interesting
variations over time. While this is easily controlled by scientists who conduct
experimental research, it can be more difficult for scientists who rely on trace
data or naturally occurring data. However, where possible, we recommend
political scientists work to expand efforts for improving the temporal
granularity of the data. One possible means of achieving this may be through
collaborative efforts like the Open Event Data Alliance [@schrodt2014three].

This methodology should be of great interest to social scientists investigating
social mobilization. It has potential applications for gender studies and 
the #MeToo movement and for the relationship between social protest and repression
during the Black Lives Matter protests. Fundamentally, this methodological
approach can help us better understand the human social mechanisms underlying
social media and its influence on human behavior in real life. Given that many
governments enact repressive measures when they determine that collective action
by citizens is imminent, this methodology should help to improve the predictive
power of using social media data in estimating repressive behavior. If social
media messages are increasing in cohesion, this may represent a threat to civil
and political liberties of citizens engaging in collective action.

## Conclusion
Considering the pervasiveness of social media, it is unsurprising that social
activists and protesters have utilized these platforms to support real-world
mobilization. Here, we quantitatively explored the connection between online
social cohesion and real-world events. In addition to introducing political
scientists to a family of powerful nonlinear dynamics analyses, our results have
serious implications for monitoring global movements in consistent across other
social uprisings. While state-run media in regions of authoritarian rule are
often biased and serve as propaganda for the leader and central government, the
self-organized nature of Twitter provides a new avenue to gain insight into
truths about uprisings and social mobilization. By conceptualizing Twitter and
real-world events as two elements of an inextricably linked dynamical system,
the levels of social cohesion within social media platforms could be monitored
for fluctuations indicating shifts in power and peace. Social media mobilization
may thus help alert the world to citizens' actions and reactions that are
suppressed by government regimes, providing insight into events that may not
otherwise be broadcast and allowing for faster intervention.

# Acknowledgements

This work was completed in part as coursework done by lead author M. Chiovaro in
the ``Applications of Nonlinear Time Series Analysis'' graduate course at the
University of Connecticut (co-taught by coauthor A. Paxton and Steven J.
Harrison).

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup